---
title: "PS270 Final Project"
author: "Evan Olesker"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



### Introduction

Throughout history, there has always been a plethora of conflict in the Middle East. In recent times, there are no exceptions. While I have followed recent activities closely, I have always wondered how these wars make the citizens think. So that's how I came up with my research question: Does high-intensity conflict increase the public's desire for religiously influenced government, or does it push them toward more secular, liberal governance? I originally thought that high-conflict zones would lead their citizens to desire a more religious and conservative influence in government. This is an important question because it helps society understand whether conflict leads to more religious conservatism in government. 


### Data Section

This project uses a cross-sectional research design to analyze the relationship between conflict and political preferences across seven different countries in the Middle East and North African region (MENA). I merged the data from two primary sources: the Armed Conflict Location & Event Data Project (ACLED) and the Arab Barometer (Wave VIII) for public opinion data.

The independent variable in my project is conflict intensity. This is measured by the total fatalities per country recorded in the ACLED dataset. The total fatalities will provide the necessary variation to compare stable countries versus countries that are in high-conflict environments. The dependent variable in my project is the political ideology. This measures public opinion towards the role of religion of government. I want to say that most countries where there is a high level of religion in government tend to be more conservative. The data was measured in a public opinion survey, and taken on a scale of one to four. One is "Strongly Agree" for religion in government, and four is "Strongly Disagree".


I believe that there will be a positive correlation between conflict and a desire for a more conservative and religiously influenced government. However, due to the way that the survey was conducted, my hypothesis will be proven if there is a negative correlation. A negative correlation will be if a higher number of fatalities leads to a lower survey score. 


```{r}
library(tidyverse)
```


```{r}
acled_data <- data.frame(
  country = c("Jordan", "Lebanon", "Palestine", "Yemen", "Iraq", "Libya", "Tunisia"),
  fatalities = c(5, 120, 30000, 1500, 400, 150, 20)
)

acled_data
```

```{r}
ab_data <- data.frame(
  country = c("Jordan", "Lebanon", "West Bank and Gaza", "Yemen", "Iraq", "Libya", "Tunisia"),
  original_score = c(2.1, 1.8, 3.4, 3.1, 2.5, 2.8, 1.9)
)

ab_data
```

```{r}
acled_clean <- acled_data %>%
  mutate(country = recode(country, "Palestine" = "West Bank and Gaza"))

acled_clean
```


```{r}
merged_data <- inner_join(acled_clean, ab_data, by = "country")

merged_data
```

Lower the number for original score = higher the desire for religious influence desired in the government. 

```{r}
ggplot(merged_data, aes(x = fatalities, y = original_score)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Conflict and Political Views",
    x = "Total Fatalities",
    y = "Average Survey Score (1=High, 4=Low)"
  )
```


```{r}
library(knitr)
ggplot(merged_data, aes(x = reorder(country, original_score), y = original_score)) +
  geom_col(fill = "steelblue") +
  theme_minimal() +
  labs(
    title = "Opposition to Religious Influence by Country",
    subtitle = "Higher scores indicate more disagreement with religious influence",
    x = "Country",
    y = "Average Survey Score (1-4)"
  )
```


### Results

```{r}
library(broom)

model1 <- lm(original_score ~ fatalities, data = merged_data)

model_results <- tidy(model1)

kable(model_results, 
      digits = 3, 
      caption = "Table 1: Regression Analysis of Conflict on Political Views",
      col.names = c("Term", "Estimate", "Std. Error", "Statistic", "P-Value"))

ggplot(merged_data, aes(x = fatalities, y = original_score)) +
  geom_point(size = 3, color = "white") +
  geom_smooth(method = "lm", color = "green", se = FALSE) +
  theme_minimal() +
  labs(
    title = "Conflict Intensity vs. Religious Support",
    subtitle = "Red line represents the OLS regression trend",
    x = "Total Fatalities (ACLED)",
    y = "Survey Score (1=Agree, 4=Disagree)"
  )
```


  In the data set that I used, the survey asked the question of "To what extent do you agree or disagree with the following statement: Religious leaders should have influence over the decisions of government". The answer would then be given on a scale from 1-4. 1 was strongly agree, and 4 was strongly disagree. Originally, I thought that areas where there was more conflict would want a more conservative and more religiously influenced government. But as I have noticed from my early research, this seems to be the opposite. In my second visualization, the table shows that the more casualties a country has, the more they disagree with having religion involved in government. Gaza and the West Bank, who has the highest casualties, also has the strongest disagreement rating from the survey. This sticks out to me because Hamas, a terrorist organization built on the foundation of religion, is currently managing that area. As I move forward in my research process, I look forward to diving into this further and looking at more data in order to see if this holds up.
