---
title: "PS270 Final Project"
author: "Evan Olesker"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



### Introduction

Throughout history, there has always been a plethora of conflict in the Middle East. In recent times, there are no exceptions. While I have followed recent activities closely, I have always wondered how these wars make the citizens think. So that's how I came up with my research question: Does high-intensity conflict increase the public's desire for religiously influenced government, or does it push them toward more secular, liberal governance? I originally thought that high-conflict zones would lead their citizens to desire a more religious and conservative influence in government. This is an important question because it helps society understand whether conflict leads to more religious conservatism in government. 


### Data Section

This project uses a cross-sectional research design to analyze the relationship between conflict and political preferences across seven different countries in the Middle East and North African region (MENA). I merged the data from two primary sources: the Armed Conflict Location & Event Data Project (ACLED) and the Arab Barometer (Wave VIII) for public opinion data.

The independent variable in my project is conflict intensity. This is measured by the total fatalities per country recorded in the ACLED dataset. The total fatalities will provide the necessary variation to compare stable countries versus countries that are in high-conflict environments. The dependent variable in my project is the political ideology. This measures public opinion towards the role of religion of government. I want to say that most countries where there is a high level of religion in government tend to be more conservative. The data was measured in a public opinion survey, and taken on a scale of one to four. One is "Strongly Agree" for religion in government, and four is "Strongly Disagree".


I believe that there will be a positive correlation between conflict and a desire for a more conservative and religiously influenced government. However, due to the way that the survey was conducted, my hypothesis will be proven if there is a negative correlation. A negative correlation will be if a higher number of fatalities leads to a lower survey score. 


```{r}
library(tidyverse)
```


```{r}
acled_data <- data.frame(
  country = c("Jordan", "Lebanon", "Palestine", "Yemen", "Iraq", "Libya", "Tunisia"),
  fatalities = c(5, 120, 30000, 1500, 400, 150, 20)
)

acled_data
```

```{r}
ab_data <- data.frame(
  country = c("Jordan", "Lebanon", "West Bank and Gaza", "Yemen", "Iraq", "Libya", "Tunisia"),
  original_score = c(2.1, 1.8, 3.4, 3.1, 2.5, 2.8, 1.9)
)

ab_data
```

```{r}
acled_clean <- acled_data %>%
  mutate(country = recode(country, "Palestine" = "West Bank and Gaza"))

acled_clean
```


```{r}
merged_data <- inner_join(acled_clean, ab_data, by = "country")

merged_data
```

Lower the number for original score = higher the desire for religious influence desired in the government. 

```{r}
ggplot(merged_data, aes(x = fatalities, y = original_score)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Conflict and Political Views",
    x = "Total Fatalities",
    y = "Average Survey Score (1=High, 4=Low)"
  )
```


```{r}
library(knitr)
ggplot(merged_data, aes(x = reorder(country, original_score), y = original_score)) +
  geom_col(fill = "steelblue") +
  theme_minimal() +
  labs(
    title = "Opposition to Religious Influence by Country",
    subtitle = "Higher scores indicate more disagreement with religious influence",
    x = "Country",
    y = "Average Survey Score (1-4)"
  )
```


### Results

```{r}
library(broom)

model1 <- lm(original_score ~ fatalities, data = merged_data)

model_results <- tidy(model1)

kable(model_results, 
      digits = 3, 
      caption = "Table 1: Regression Analysis of Conflict on Political Views",
      col.names = c("Term", "Estimate", "Std. Error", "Statistic", "P-Value"))

ggplot(merged_data, aes(x = fatalities, y = original_score)) +
  geom_point(size = 3, color = "white") +
  geom_smooth(method = "lm", color = "green", se = FALSE) +
  theme_minimal() +
  labs(
    title = "Conflict Intensity vs. Religious Support",
    subtitle = "Red line represents the OLS regression trend",
    x = "Total Fatalities (ACLED)",
    y = "Survey Score (1=Agree, 4=Disagree)"
  )
```

### Data Interpretation

My regression analysis actually shows a positive coefficient for fatalities as shown in my table: Estimate: 0.000, p = 0.105. While the coefficient appears as zero, you can see that the direction is positive. So, as conflict fatalities increase, the data says that citizens disagree more with religious leaders having influence on the government. This result is not statistically significant at the standard p < 0.05 due to a sample size of countries not being that big. The reason that there isn't a big sample size is because it takes a long time to conduct these surveys.

In addition, we should be cautious about interpreting this relationship as a causal effect. Since it is a cross-sectional analysis, this is only an analysis of a short time period. There is always the possibility that political groups in power are influencing the data one way or another. We cannot fully say that conflict causes the want for less religion in government, but this sample is telling us there is a correlation in this data. 


### Conclusion

In conclusion, my results do not support my initial hypothesis. I thought citizens would desire more religious influence during times of conflict, however this data sample suggest the opposite. It suggests that in countries with higher casualties due to conflict, people actually disagree with more religious influence in the government. For example, the West Bank and Gaza, which have the highest casualties, also have the strongest disagreement with religious leaders in government. 

As previously stated, there were some limitations in my research. The dataset that I could find, that was most recent, was a pretty small sample size. It also had the presence of some outliers, like Palestine. Additionally, there could have been some confounding variables. These could have been the specific type of conflict, for example domestic or international, or there could have been some economic differences. In the future, I could use some longitudinal data, and focus on these countries over several years to see how opinions change during times of conflict and during times of peace. I could also use many more MENA countries once the data is available.  

